Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2025-02-09 paper/main.tex
Updates

% Our approach combines implied Conditional Independence (CI) testing with a
% measure of conditional association between variables to rank violations
% of implied CIs in a given DAG. This ranking helps prioritize modifications to
% the DAG to improve its fit to the data while utilizing expert knowledge to
% decide the orientation of new edges.

@@

\section{Introduction}
Understanding cause-and-effect relationships between variables is a fundamental
objective in many scientific fields. These relationships reveal the mechanisms
behind observed phenomena and guide effective interventions or policy
decisions. Causal discovery methods aim to discover such relationships
among random variables using observational data. Approaches to causal discovery
have been developed within both the Directed Acyclic Graphs (DAGs) and
Structural Equation Models (SEMs) frameworks, each adapting a different
approach.

In the DAG literature, the primary focus has been on developing automated
algorithms to learn causal structures from datasets. These efforts have led to
numerous causal discovery algorithms, such as constraint-based methods like PC
algorithm \citep{Spirtes2001} and Fast Causal Inference \citep{Spirtes2000}),
score-based methods such as Hill-Climb Search and Greedy Equivalence Search
\citep{Chickering2002}, and continuous optimization-based methods like NO TEARS
\citep{Zheng2018} and DAGMA \citep{Bello2022}. While DAG-based methods focus on
automated discovery, SEM-based methods emphasize expert driven model
specification. This includes tools to assist researchers in manually
constructing models, enabling them to incorporate their domain knowledge in the
model building process. Researchers typically begin with an initial model based
on their domain knowledge and then use these tools to guide modifications that
improve the model's fit to data. This process is commonly known as
Specification Search \citep{Long1983} and uses method such as modification
indices, and Wald-based tests \citep{Marcoulides2018}.

>>>>

@@

\section{Introduction}
Understanding cause-and-effect relationships between variables is a fundamental
objective in many scientific fields. These relationships reveal the mechanisms
behind observed phenomena and guide effective interventions or policy
decisions. Causal discovery methods aim to discover such relationships
among random variables using observational data.

% Approaches to causal discovery
% have been developed within both the Directed Acyclic Graphs (DAGs) and
% Structural Equation Models (SEMs) frameworks, each adapting a different
% approach.

In the DAG literature, the primary focus has been on developing automated
algorithms to learn causal structures from datasets. These efforts have led to
numerous causal discovery algorithms, such as constraint-based methods like PC
algorithm \citep{Spirtes2001} and Fast Causal Inference \citep{Spirtes2000}),
score-based methods such as Hill-Climb Search and Greedy Equivalence Search
\citep{Chickering2002}, and continuous optimization-based methods like NO TEARS
\citep{Zheng2018} and DAGMA \citep{Bello2022}.

% While DAG-based methods focus on
% automated discovery, SEM-based methods emphasize expert driven model
% specification. This includes tools to assist researchers in manually
% constructing models, enabling them to incorporate their domain knowledge in the
% model building process. Researchers typically begin with an initial model based
% on their domain knowledge and then use these tools to guide modifications that
% improve the model's fit to data. This process is commonly known as
% Specification Search \citep{Long1983} and uses method such as modification
% indices, and Wald-based tests \citep{Marcoulides2018}.
